---
title: "04 - Facility Location Algorithm"
author: "Alec Stashevsky"
date: "July 3, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 10) # need at least 8 decimal places for coordinates
library(readxl)
library(openxlsx)
library(tidyverse)
library(data.table)
library(geosphere)
library(maps)
library(sp)
rm(list = ls()) # Clean R Environment
```

## Import Geodistance Data from 03
```{r Import}
path.in <- "C:/Users/alec.stashevsky/OneDrive - Cadmus/Documents/GAP Climate Research/GeoData/APA Geodistance Data.xlsx"

nyc <- read_excel(path.in, sheet = "NYC 2018") %>% setDT()
sf <- read_excel(path.in, sheet = "SF 2019") %>% setDT()
phl <- read_excel(path.in, sheet = "PHL 2020") %>% setDT()
conventions <- read_xlsx(path.in, sheet = "Convention Centers") %>%
  setDT()
```

## Facility Location Algorithm

Latitudes range from [-90, 90] and longitudes ranges from [-180, 180]. Here we create an algorithm to identify the optimal conference location based on the APA Annual Meetings geographic makeup of attendees.

### NYC 2018
```{r Facility Location Algorithm NYC}
# Sample 30000 coordinates for initial whole-earth grid
## This number can be increased to augment accuracy
set.seed(32)
coordinate.sample.nyc <- as.data.table(cbind(
  lon = round(runif(30000, -180, 180), 8),
  lat = round(runif(30000, -90, 90), 8)))

# Calculate geodesic distance for each coordinate
## Again, we are using the WGS84 ellipsoidal distance
nyc.dist.log <- rep(NA, nrow(coordinate.sample.nyc))
for (i in 1:nrow(coordinate.sample.nyc)){
  nyc.dist.log[i] <- sum(nyc$Frequency * # Weight by frequency
                           distGeo(nyc[, c('lon', 'lat')],
                                   c(coordinate.sample.nyc$lon[i],
                                     coordinate.sample.nyc$lat[i])),
                         na.rm = TRUE)
}

coordinate.rank.nyc <- arrange(cbind(coordinate.sample.nyc, nyc.dist.log),
                               nyc.dist.log)
coordinate.rank.nyc <- cbind(coordinate.rank.nyc,
                             rank = seq(1:nrow(coordinate.rank.nyc)))

# Create grid of confidence from 10 coordinates with the least geo-distance
grid.nyc <- as.matrix(rbind(
  c(min(coordinate.rank.nyc$lon[1:10]),
    min(coordinate.rank.nyc$lat[1:10])),
  c(max(coordinate.rank.nyc$lon[1:10]),
    max(coordinate.rank.nyc$lat[1:10]))))
```

### SF 2019
```{r Facility Location Algorithm SF}
# Sample 30000 coordinates for initial whole-earth grid
## This number can be increased to augment accuracy
set.seed(656)
coordinate.sample.sf <- as.data.table(cbind(
  lon = round(runif(30000, -180, 180),8),
  lat = round(runif(30000, -90, 90), 8)))

# Calculate geodesic distance for each coordinate
## Again, we are using the WGS84 ellipsoidal distance
sf.dist.log <- rep(NA, nrow(coordinate.sample.sf))
for (i in 1:nrow(coordinate.sample.sf)){
  sf.dist.log[i] <- sum(sf$Frequency * # Weight by frequency
                          distGeo(sf[, c('lon', 'lat')],
                                  c(coordinate.sample.sf$lon[i],
                                    coordinate.sample.sf$lat[i])), na.rm = TRUE)
}

coordinate.rank.sf <- arrange(cbind(coordinate.sample.sf, sf.dist.log),
                              sf.dist.log)
coordinate.rank.sf <- cbind(coordinate.rank.sf,
                            rank = seq(1:nrow(coordinate.rank.sf)))

# Create grid of confidence from 10 coordinates with the least geo-distance
grid.sf <- as.matrix(rbind(
  c(min(coordinate.rank.sf$lon[1:10]),
    min(coordinate.rank.sf$lat[1:10])),
  c(max(coordinate.rank.sf$lon[1:10]),
    max(coordinate.rank.sf$lat[1:10]))))
```

### PHL 2020
```{r Facility Location Algorithm PHL}
# Sample 30000 coordinates for initial whole-earth grid
## This number can be increased to augment accuracy
set.seed(321)
coordinate.sample.phl <- as.data.table(cbind(
  lon = round(runif(30000, -180, 180), 8),
  lat = round(runif(30000, -90, 90), 8)))

# Calculate geodesic distance for each coordinate
## Again, we are using the WGS84 ellipsoidal distance
phl.dist.log <- rep(NA, nrow(coordinate.sample.phl))
for (i in 1:nrow(coordinate.sample.phl)){
  phl.dist.log[i] <- sum(phl$Frequency * # Weight by frequency
                           distGeo(phl[, c('lon', 'lat')],
                                   c(coordinate.sample.phl$lon[i],
                                     coordinate.sample.phl$lat[i])),
                         na.rm = TRUE)
}

coordinate.rank.phl <- arrange(cbind(coordinate.sample.phl, phl.dist.log),
                               phl.dist.log)

coordinate.rank.phl <- cbind(coordinate.rank.phl,
                             rank = seq(1:nrow(coordinate.rank.phl)))

# Create grid of confidence from 10 coordinates with the least geo-distance
grid.phl <- as.matrix(rbind(
  c(min(coordinate.rank.phl$lon[1:10]),
    min(coordinate.rank.phl$lat[1:10])),
  c(max(coordinate.rank.phl$lon[1:10]),
    max(coordinate.rank.phl$lat[1:10]))))
```


## Facility Location Plot

These plots display a region of relative confidence which minimizes the distance APA Annual Meeting attendees will travel. Provided the attendee makeup is similar to that of of each We can locate the convention center with necessary floor space and capacity closest to centroid of our region of confidence. The coordinates for this point are (lon = -77.35697915, lat = 42.02497199).

### NYC 2018

```{r Facility Location Plot NYC}
# Initialize write to pdf
pdf(file = "APA Facility Location Simulation.pdf")

# Plot 15 best sample coordinates to create region of confidence
map("state", col="grey20", fill=TRUE, bg="black", lwd=0.1,
    xlim = c(grid.nyc[1,1] - 3, grid.nyc[2,1] + 4),
    ylim = c(grid.nyc[1,2] - 3, grid.nyc[2,2] + 4))

points(x = coordinate.rank.nyc$lon[1:10],
       y = coordinate.rank.nyc$lat[1:10],
       col = "orange red",
       pch = 10, cex = 2.8)

text(x = coordinate.rank.nyc$lon[1:10],
     y = coordinate.rank.nyc$lat[1:10],
     labels = coordinate.rank.nyc$rank[1:10],
     col = "white", cex = 0.8)

# Isolate maximum bounding polygon with best 15 approximations
target <- c(10, 4, 6, 9, 5, 7, 8)
poly.ordering.nyc <- as.matrix(rbind(
  coordinate.rank.nyc[target[1]],
  coordinate.rank.nyc[target[2]],  
  coordinate.rank.nyc[target[3]],
  coordinate.rank.nyc[target[4]],   
  coordinate.rank.nyc[target[5]],
  coordinate.rank.nyc[target[6]],
  coordinate.rank.nyc[target[7]]))

# Identify the centroid/geometric median/center of mass
## This is actually calculated by using Euclidean distance of Mercator projection
region.nyc <- makePoly(poly.ordering.nyc[,1:2], sp = TRUE)
center.nyc <- centroid(region.nyc)

polygon(poly.ordering.nyc, col = 
          rgb(red = 1, green = 140/255, blue = 0, alpha = 0.25),
        border = "orange")
# Add center point
points(x = center.nyc[1], y= center.nyc[2], col = "gold", pch = 3)

# Add plot aesthetics
title(main = "APA Facility Location Simulation - NYC 2018",
      col.main = "white")
```

### SF 2019
```{r Facility Location Plot SF}
# Plot 15 best sample coordinates to create region of confidence
map("state", col="grey20", fill=TRUE, bg="black", lwd=0.1,
    xlim = c(grid.sf[1,1] - 3, grid.sf[2,1] + 4),
    ylim = c(grid.sf[1,2] - 3, grid.sf[2,2] + 4))

points(x = coordinate.rank.sf$lon[1:10],
       y = coordinate.rank.sf$lat[1:10],
       col = "cyan",
       pch = 10, cex = 2.8)

text(x = coordinate.rank.sf$lon[1:10],
     y = coordinate.rank.sf$lat[1:10],
     labels = coordinate.rank.sf$rank[1:10],
     col = "white", cex = 0.8)

# Isolate maximum bounding polygon with best 15 approximations
target <- c(10, 7, 8, 6, 9, 4, 5)
poly.ordering.sf <- as.matrix(rbind(
  coordinate.rank.sf[target[1]],
  coordinate.rank.sf[target[2]],
  coordinate.rank.sf[target[3]],
  coordinate.rank.sf[target[4]],
  coordinate.rank.sf[target[5]],
  coordinate.rank.sf[target[6]],
  coordinate.rank.sf[target[7]]))

# Identify the centroid/geometric median/center of mass
## This is actually calculated by using Euclidean distance of Mercator projection
region.sf <- makePoly(poly.ordering.sf[,1:2], sp = TRUE)
center.sf <- centroid(region.sf)


polygon(poly.ordering.sf, col = 
          rgb(red = 0, green = 1, blue = 1, alpha = 0.25),
        border = "cyan")
# Add center point
points(x = center.sf[1], y= center.sf[2], col = "gold", pch = 3)

# Add plot aesthetics
title(main = "APA Facility Location Simulation - SF 2019",
      col.main = "white")
```

### PHL 2020
```{r Facility Location Plot PHL}
# Plot 15 best sample coordinates to create region of confidence
map("state", col="grey20", fill=TRUE, bg="black", lwd=0.1,
    xlim = c(grid.phl[1,1] - 3, grid.phl[2,1] + 4),
    ylim = c(grid.phl[1,2] - 3, grid.phl[2,2] + 4))

points(x = coordinate.rank.phl$lon[1:10],
       y = coordinate.rank.phl$lat[1:10],
       col = "purple",
       pch = 10, cex = 2.8)

text(x = coordinate.rank.phl$lon[1:10],
     y = coordinate.rank.phl$lat[1:10],
     labels = coordinate.rank.phl$rank[1:10],
     col = "white", cex = 0.8)

# Isolate maximum bounding polygon with best 15 approximations
target <- c(10, 9, 5, 3, 7, 6, 8)
poly.ordering.phl <- as.matrix(rbind(
  coordinate.rank.phl[target[1]],
  coordinate.rank.phl[target[2]],
  coordinate.rank.phl[target[3]],
  coordinate.rank.phl[target[4]],
  coordinate.rank.phl[target[5]],
  coordinate.rank.phl[target[6]],
  coordinate.rank.phl[target[7]]))

# Identify the centroid / geometric median/center of mass
## This is actually calculated by using Euclidean distance of Mercator projection
region.phl <- makePoly(poly.ordering.phl[,1:2], sp = TRUE)
center.phl <- centroid(region.phl)

polygon(poly.ordering.phl, col = 
          rgb(red = 1, green = 107/255, blue = 180/255, alpha = 0.25),
        border = "purple")
# Add center point
points(x = center.phl[1], y= center.phl[2], col = "gold", pch = 3)

# Add plot aesthetics
title(main = "APA Facility Location Simulation - PHL 2020 (Pre-cancellation)",
      col.main = "white")
```


```{r Combine Plots}

map("state", col="grey20", fill=TRUE, bg="black", lwd=0.1,
    xlim = c(grid.nyc[1,1] - 10, grid.nyc[2,1] + 6),
    ylim = c(grid.nyc[1,2] - 15, grid.nyc[2,2] + 10))

polygon(poly.ordering.nyc, col = 
          rgb(red = 1, green = 140/255, blue = 0, alpha = 0.25),
        border = "orange")
# Add center point
points(x = center.nyc[1], y= center.nyc[2], col = "orange", pch = 3)

polygon(poly.ordering.sf, col = 
          rgb(red = 0, green = 1, blue = 1, alpha = 0.25),
        border = "cyan")
# Add center point
points(x = center.sf[1], y= center.sf[2], col = "cyan", pch = 3)

polygon(poly.ordering.phl, col = 
          rgb(red = 1, green = 107/255, blue = 180/255, alpha = 0.25),
        border = "purple")
# Add center point
points(x = center.phl[1], y= center.phl[2], col = "purple", pch = 3)

# Add plot aesthetics
title(main = "APA Facility Location Simulation",
      col.main = "white")

legend("bottom", legend=c("NYC 2018", "SF 2019", "PHL 2020 (Pre-cancellation)"),
       col=c("orange red", "cyan", "purple"), lty = 1, lwd = 1.2,  cex=0.75,
       bg = "black",
       text.col = "white")

dev.off()
```