---
title: "01 - Regional Data Prep"
author: "Alec Stashevsky"
date: "12/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(openxlsx)
library(tidyverse)
library(data.table)
library(geosphere)
library(ggmap)
library(sf)
library(spData)
library(usdata)
library(rgdal)
rm(list = ls()) # Clean R Environment
```

## Import Geodistance + Airport Data
```{r Import}
path.in <- "C:/Users/Alec/Documents/GAP Research/GAP Climate Research/GeoData/APA Geodistance + Airport Data.xlsx"

nyc <- read_excel(path.in, sheet = "NYC 2018") %>% setDT()
sf <- read_excel(path.in, sheet = "SF 2019") %>% setDT()
phl <- read_excel(path.in, sheet = "PHL 2020") %>% setDT()

# Import APA Regions
path.regions <- "C:/Users/Alec/Documents/GAP Research/GAP Climate Research/Data/APA Regions.xlsx"
regions <- read_excel(path.regions) %>% setDT()

# Import list of convention centers
convention.import <- read_excel("C:/Users/Alec/Documents/GAP Research/GAP Climate Research/GeoData/Convention Centers.xlsx", range = "A1:E16")

# Import Global Airport Database
path.airports <- "C:/Users/Alec/Documents/GAP Research/GAP Climate Research/GlobalAirportDatabase/GlobalAirportDatabase.txt"

# Import that data
airports.raw <- read.delim(path.airports, header = FALSE, sep = ":") %>% setDT()

# Drop extraneous columns
airports <- airports.raw[, c(1:5, 14:16)]

# Set column names
setnames(airports, names(airports), c("ICAO", "IATA", "Airport Name",
         "City", "Country", "Altitude", "Latitude", "Longitude"))


```


We only need to add one new regional meeting. Others will take place in previous Annual Meeting Locations.
Areas:
  1. Toronto
  2. New York City
  3. Philadelphia
  4. Chicago
  5. New Orleans
  6. Los Angeles
  7. Seattle (New)

## Remove old feilds from main analysis
```{r Clean Feilds}
# List of IATA codes to remove from data
iata.rm <- c("SFO", "DFW", "DCA", "YUL", "MIA", "SAN", "ATL", "HNL")

# Remove columns w/ un-needed fields 
for (i in 1:length(iata.rm)) {
  
  nyc[, grep(iata.rm[i], names(nyc), ignore.case = TRUE) := NULL]
  sf[, grep(iata.rm[i], names(sf), ignore.case = TRUE) := NULL]
  phl[, grep(iata.rm[i], names(phl), ignore.case = TRUE) := NULL]
  
}

# Fix unabbreviated state
nyc[STATE == "NORTH CAROLINA"]$STATE <- "NC"
sf[STATE == "OHIO"]$STATE <- "OH"

# Fix US/Canada mixing
nyc[COUNTRY == "United States" & STATE == "ON"]$STATE <- "CA"
nyc[COUNTRY == "United States" & STATE == "BC"]$COUNTRY <- "Canada"
sf[COUNTRY == "United States" & STATE == "ON"]$COUNTRY <- "Canada"
sf[COUNTRY == "Canada" & STATE == "NY"]$COUNTRY <- "United States"
```


## Add Seattle Convention Center
```{r Seattle Convention}
# Geocode Seattle convention center
conventions.geo <- mutate_geocode(convention.import, location = location)

# Get destination airports
airports.dests <- rbind(airports[IATA == 'PHL'],
                        airports[IATA == 'SFO'],
                        airports[IATA == 'JFK'],
                        airports[IATA == 'ORD'],
                        airports[IATA == 'MSY'],
                        airports[IATA == 'YYZ'],
                        airports[IATA == 'LAX'],
                        airports[IATA == 'DFW'],
                        airports[IATA == 'DCA'],
                        airports[IATA == 'YUL'],
                        airports[IATA == 'MIA'],
                        airports[IATA == 'SAN'],
                        airports[IATA == 'ATL'],
                        airports[IATA == 'HNL'],
                        airports[IATA == 'SEA']
                        )

# Attach onto the convention center sheet
conventions <- data.table(conventions.geo, airports.dests)

# Export 
write.xlsx(conventions, "C:/Users/Alec/Documents/GAP Research/GAP Climate Research/GeoData/APA Convention Center GeoData.xlsx")
```


## Geodistance to Seattle
```{r SEA Geodistance}
# Get convention center coordinates
sea.dest <- as.numeric(conventions[IATA == "SEA", .(lon, lat)])

# Calculate Geodistance in meters - with ellipsoidal method
nyc.gdist.SEA <- distGeo(nyc[, .(lon, lat)], sea.dest)
sf.gdist.SEA <- distGeo(sf[, .(lon, lat)], sea.dest)
phl.gdist.SEA <- distGeo(phl[, .(lon, lat)], sea.dest)

# Merge to data
nyc <- data.table(nyc, "gdist.SEA" = nyc.gdist.SEA)
sf <- data.table(sf, "gdist.SEA" = sf.gdist.SEA)
phl <- data.table(phl, "gdist.SEA" = phl.gdist.SEA)
```


## Driving distance to SEA indicator
```{r Driving Distance}
# Create indicator variable for driving if less than 400 km away from meeting
nyc[, `:=`(drive.SEA = ifelse(gdist.SEA <= 400000, 1, 0))]
sf[, `:=`(drive.SEA = ifelse(gdist.SEA <= 400000, 1, 0))]
phl[, `:=`(drive.SEA = ifelse(gdist.SEA <= 400000, 1, 0))]
```


## Recode State by coordinates
```{r Recode US States}
coord_to_state <- function(pointsDF,
                           states = spData::us_states,
                           name_col = "NAME") {
  
    ## Convert points data.frame to an sf POINTS object
    pts <- st_as_sf(pointsDF, coords = 1:2, crs = 4326)

    ## Transform spatial data to some planar coordinate system
    ## (e.g. Web Mercator) as required for geometric operations
    states <- st_transform(states, crs = 3857)
    pts <- st_transform(pts, crs = 3857)

    ## Find names of state (if any) intersected by each point
    state_names <- states[[name_col]]
    ii <- as.integer(st_intersects(pts, states))
    output <- state_names[ii]
}

test <- coord_to_state(nyc[, .(lon, lat)])

for (i in 1:ncol(regions)) {
  print(state2abbr(regions$`AREA 1`))
}


```


## Clean US and CA States
```{r Reginal Indicators}
# Clean any Canadian's w/o Provence info
nyc[COUNTRY =="Canada" & is.na(STATE) & City =="MONTREAL"]$STATE <- "QC"
nyc[COUNTRY =="Canada" & is.na(STATE) & City =="WINNIPEG"]$STATE <- "MB"
nyc[COUNTRY =="Canada" & is.na(STATE) & City =="TERRACE"]$STATE <- "BC"
nyc[COUNTRY =="Canada" & is.na(STATE) &
      City %in% c("FORT MCMURRAY", "GRANDE PRAIRIE")]$STATE <- "AB"

nyc[COUNTRY =="Canada" & STATE == "CO"]$STATE <- "ON"

sf[COUNTRY =="Canada" & is.na(STATE) & City == "KAMLOOPS"]$STATE <- "BC"
sf[COUNTRY =="Canada" & is.na(STATE) &
     City %in% c("LA GRANDE RIVIERE", "GATINEAU")]$STATE <- "QC"

# United States
nyc[COUNTRY =="United States" & is.na(STATE) & City == "EVERETT"]$STATE <- "WA"
nyc[COUNTRY =="United States" & is.na(STATE) & City == "TULSA"]$STATE <- "OK"
nyc[COUNTRY =="United States" & is.na(STATE) & City == "FORT DRUM"]$STATE <- "NY"

sf[COUNTRY =="United States" & is.na(STATE) & City == "CHARLOTTE"]$STATE <- "NC"
sf[COUNTRY =="United States" & is.na(STATE) & City == "TULSA"]$STATE <- "OK"
sf[COUNTRY =="United States" & is.na(STATE) & City == "EVERETT"]$STATE <- "WA"

# Fix Spain location coded as United States
sf[COUNTRY =="United States" & is.na(STATE) & City == "MADRID"]$COUNTRY <- "Spain"
```


## APA Regional Indicators
```{r Regional Indicators}
# Create area lists for US
area1 <- state2abbr(regions$`AREA 1`)[!is.na(state2abbr(regions$`AREA 1`))]
area1 <- area1[area1 != "NV"] # remove mis-coded state
area2 <- state2abbr(regions$`AREA 2`)[!is.na(state2abbr(regions$`AREA 2`))]
area3 <- state2abbr(regions$`AREA 3`)[!is.na(state2abbr(regions$`AREA 3`))]
area3[area3 == "WA"] <- "DC" # recode Washington DC instead of WA
area4 <- state2abbr(regions$`AREA 4`)[!is.na(state2abbr(regions$`AREA 4`))]
area5 <- state2abbr(regions$`AREA 5`)[!is.na(state2abbr(regions$`AREA 5`))]
area5 <- c(area5, "PR", "AS", "AE") # Add Puerto Rico and leftover military locations
area6 <- state2abbr(regions$`AREA 6`)[!is.na(state2abbr(regions$`AREA 6`))]
area7 <- state2abbr(regions$`AREA 7`)[!is.na(state2abbr(regions$`AREA 7`))]
area7 <- head(area7, length(area7) - 1) # remove extra NV from Canadian Provence

# Create area lists for Canada
area1.ca <- c("QC", "ON", "NS", "NB", "NL", "PE") # Eastern Canada
area7.ca <- c("AB", "BC", "MB", "SK", "YT", "NT", "NU") # Western Canada

# New York Base
nyc[COUNTRY == "United States", `:=`(APA.region = fcase(STATE %in% area1, 1,
                                                        STATE %in% area2, 2,
                                                        STATE %in% area3, 3,
                                                        STATE %in% area4, 4,
                                                        STATE %in% area5, 5,
                                                        STATE %in% area6, 6,
                                                        STATE %in% area7, 7))]

nyc[COUNTRY == "Canada", `:=`(APA.region = fcase(STATE %in% area1.ca, 1,
                                                 STATE %in% area7.ca, 7)
                              )]

# San Francisco Base
sf[COUNTRY == "United States", `:=`(APA.region = fcase(STATE %in% area1, 1,
                                                       STATE %in% area2, 2,
                                                       STATE %in% area3, 3,
                                                       STATE %in% area4, 4,
                                                       STATE %in% area5, 5,
                                                       STATE %in% area6, 6,
                                                       STATE %in% area7, 7))]

sf[COUNTRY == "Canada", `:=`(APA.region = fcase(STATE %in% area1.ca, 1,
                                                 STATE %in% area7.ca, 7)
                              )]

## Do PHL even though it was not cleaned for weird US/ Canada states
# Philadelphia Base
phl[COUNTRY == "United States", `:=`(APA.region = fcase(STATE %in% area1, 1,
                                                       STATE %in% area2, 2,
                                                       STATE %in% area3, 3,
                                                       STATE %in% area4, 4,
                                                       STATE %in% area5, 5,
                                                       STATE %in% area6, 6,
                                                       STATE %in% area7, 7))]

phl[COUNTRY == "Canada", `:=`(APA.region = fcase(STATE %in% area1.ca, 1,
                                                 STATE %in% area7.ca, 7)
                              )]



```


## Save Airport Data
```{r Export Data}
final.sets <- list("NYC 2018" = nyc,
                 "SF 2019" = sf,
                 "PHL 2020" = phl,
                 "Convention Centers" = conventions)

write.xlsx(final.sets,
           file = "C:/Users/Alec/Documents/GAP Research/GAP Climate Research/GeoData/Regional Analysis/APA Geodistance + Airport Data.xlsx")
```